We record the threats to the validity of this review and the limitations of the mapping.

\textbf{Interpretive construct validity.} The frameworks are written at the level of
outcomes and processes and do not prescribe methods. Our (method, requirement) evidence
strengths are therefore reasoned engineering interpretations, not legal compliance
determinations. We mitigated subjectivity with an adversarial second pass over every
non-blank cell and by grounding each assignment in the official clause text, but
reasonable experts may code individual cells differently. The matrix should be read as a
structured argument, not an authoritative legal crosswalk.

\textbf{Temporal validity.} All three frameworks are moving targets. EU AI Act delegated
acts and the CEN--CENELEC harmonised standards are still in progress; US executive-branch
AI policy has shifted over the review window; and ISO/IEC companion standards continue to
emerge. The mapping is a snapshot dated June~2026 and is intended to be maintained as a
living artifact (Section~\ref{sec:future}).

\textbf{Selection and search bias.} Despite a broad, multi-source search, database
coverage, English-only inclusion, and the inclusion of preprints introduce bias. The
reused agent-fairness corpus is, by construction, dense in fairness methods (M2); other
families draw on a newer, smaller verified set, which may under-represent some mature
methods. We report counts transparently rather than claiming exhaustiveness.

\textbf{Citation integrity.} Cited references were verified against arXiv abstract-page
resolution, the Consensus database, or official DOIs where resolvable, and the corpus was
de-duplicated; the verification caught and corrected a substantial fraction of initially
mis-attributed metadata. We do not claim exhaustive verification of every entry in the
released corpus, and residual risk remains for the newest preprints whose venues may change.

\textbf{Not legal advice.} This article is a technical review for evaluation
practitioners and researchers. It does not constitute legal advice, and conformity
determinations remain the responsibility of providers, deployers, notified bodies, and
competent authorities.
